Data Prep

### 2024-02-20 scripts/00.0_dataPrep.qmd
removed northern indivs from social as well as movement. Began to parallelize.

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
library(here)
```

@@

@@

-   `data/dataPrep/removed_lfr.Rda`

@@

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = here::here("data", "dataPrep", "roosts.Rda"))

# Tidy up
rm(removed_lfr)
rm(roosts)
gc()
```

@@

@@

```{r}
#| eval: false
# Remove nighttime points -------------------------------------------------
removed_nighttime <- map(removed_lfr, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

@@

# tidy up
rm(removed_nighttime)
rm(removed_lfr)
rm(season_names)
gc()
```

-   `data/dataPrep/removed_nighttime.Rda`
-   `data/dataPrep/season_names.Rda`

```{r}
load(here::here("data", "dataPrep", "removed_nighttime.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

```{r}
#| eval: false
# Transform to SF object, so we can get centroids
seasons_sf <- map(removed_nighttime, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

## Remove individuals not in the south
before <- removed_nighttime
removed_northern <- map2(.x = removed_nighttime, 
                         .y = southern_indivs, ~.x 
                         %>% filter(Nili_id %in% .y))
after <- removed_northern

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
df <- data.frame(season = season_names, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
df

save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))

# tidy up
rm(removed_nighttime)
rm(season_names)
rm(seasons_sf)
rm(centroids)
rm(southern_indivs)
rm(before)
rm(after)
rm(removed_northern)
rm(df)
rm(afterIndivs)
rm(afterRows)
rm(beforeIndivs)
rm(beforeRows)
gc()
```

>>>>

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
library(here)
library(furrr)
future::plan(future::multisession, workers = 10) # this will last for the whole R session.
```

@@

```{r}
#| eval: false
# Transform to SF object, so we can get centroids
seasons_sf <- map(removed_lfr, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

## Remove individuals not in the south
removed_northern <- map2(.x = removed_lfr, 
                         .y = southern_indivs, ~.x 
                         %>% filter(Nili_id %in% .y))
save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))

# tidy up
rm(removed_lfr)
rm(season_names)
rm(seasons_sf)
rm(centroids)
rm(southern_indivs)
rm(removed_northern)
rm(df)
gc()
```

@@

@@

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- furrr::future_map(removed_northern, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"), .progress = T)
roosts <- roosts %>%
  furrr::future_map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F), .progress = T)
save(roosts, file = here::here("data", "dataPrep", "roosts.Rda"))

# Tidy up
rm(removed_northern)
rm(roosts)
gc()
```

@@

-   `data/dataPrep/removed_northern.Rda`
-   `data/dataPrep/season_names.Rda`

@@

```{r}
load(here::here("data", "dataPrep", "removed_northern.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

```{r}
#| eval: false
# Remove nighttime points -------------------------------------------------
removed_nighttime <- map(removed_northern, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

@@

# tidy up
rm(removed_nighttime)
rm(removed_northern)
rm(season_names)
gc()
```
